### 2024-06-07 products/manuscript/Manuscript.qmd
Second update with code.

---
title: "Predicting and Preventing Employee Turnover through HR Analytics"
subtitle: ""
author: Guillermo Gallardo
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)

hr_data = read.csv("C:/Users/crew_/OneDrive/Escritorio/Guille/UTSA Masters/MSDA/DA6833/PracticumII-Course/guillermogallardo-final-project/data/raw-data/IBM HR Analytics Data.csv", header = TRUE)
```

# Project Overview

The primary objective of this project is to identify the key factors contributing to employee turnover and to develop a model that can predict potential departures within a company. By anticipating these departures, we can create strategies for employee retention and enhance the overall employee experience.

I find this dataset interesting, as it aligns with my previous experience in HR Analytics. Although I've worked in this field before, I never had the opportunity to work on project where we could predict attrition. This project allows me to deep dive into the data I am interested about, while also enhancing my skills in modeling using R.

# Project Question

We have two main questions that we want to answer during this project.

-   What are the primary factors that contribute to employee attrition?

-   What strategies can be implement to enhance employee retention and improve the overall employee experience?

# Dataset

For this project, I was looking for a dataset with employee data that included a wide range of variables to explore potential relationships with attrition. Below, I will provide an overview of the dataset's origin, author, the types of variables it includes, and other relevant details.

## **IBM HR Analytics Employee Attrition**

The dataset for this project was sourced from [Kaggle.com](https://www.kaggle.com/datasets/pavansubhasht/ibm-hr-analytics-attrition-dataset) and was created by IBM Data Scientists as a fictional representation of HR data. It consists of 1,470 rows and 35 variables that cover demographic details, job characteristics, compensation, performance metrics, and employee satisfaction data. This data serves as the foundation for analyzing factors contributing to employee turnover and developing a predictive model.

## **Variables**

This file includes 35 variables, detailed below, which includes a combination of character and integer types, with some of the character variables being categorical.

```{r Variables}

str(hr_data)
```

## **Gender**

Out of the 1,470 observation in the dataset, 588 are female and 882 are male. This provides a quick snapshot of the gender distribution within the data.

```{r Gender}
table(hr_data$Gender)
```

## **Education**

During data cleaning, we'll use the data dictionary to decode each field. For instance, the education field is labeled from 1 to 5, representing different educational levels: 1 = "Below College," 2 = "College," 3 = "Bachelor's," 4 = "Master's," and 5 = "Doctorate." Notably, about 40% of the dataset holds a Bachelor's degree.

```{r Education}
table(hr_data$Education)
```

# Methods

*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*

## Exploratory/Descriptive analysis

## Statistical analysis

*Explain anything related to your statistical analyses.*